---
title: "hw5"
author: "Stephen Powers"
date: "11/8/2019"
output: github_document
---

## Problem 1 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(viridis)
library(patchwork)
```


```{r}
set.seed(10)

iris_with_missing = iris %>% 
  map_df(~replace(.x, sample(1:150, 20), NA)) %>%
  mutate(Species = as.character(Species))
```

```{r}
fill_missing = function(x) {
  if (is.numeric(x)) 
    {replace_na(x, mean(x, na.rm = TRUE))} 
  else if (is.character(x)) 
    {replace_na(x, "virginica")}}

iris_fill = map_df(iris_with_missing, fill_missing)
```

## Problem 2 

```{r}
p2_data = 
  list.files("./data/", pattern = ".csv", full.names = TRUE) %>% view
```

#### Tidy Up

```{r}
p2_tidy_data = 
  p2_data %>% map_df(read.csv) %>% 
  mutate(id = tools::file_path_sans_ext(basename(p2_data))) %>% 
  separate(id, into = c("group", "id_number"), sep = "_") %>% 
  mutate(group = recode(group, "con" = "control", "exp" = "experimental")) %>% 
  mutate(observation = row_number()) %>%
  select(observation, group, id_number, everything()) %>% 
  pivot_longer(week_1:week_8,
    names_to = "week",
    values_to = "value",
    names_prefix = "week_") %>% view
```

#### Spaghetti Plot

```{r}
p2_tidy_data %>% 
  ggplot(aes(x = week, y = value, color = group, group = observation)) +
  geom_line() +
  scale_color_viridis(discrete = TRUE) +
  theme_minimal() +
  ggtitle("Observations on each subject over time by group") +
  xlab("Week") +
  ylab("Values") 

```

Based on the plot above, it appears the values are higher in the experimental group than in the control group. Also, it appears that the experimental groups' values increase over time, but the control groups do not.

## Problem 3 

#### Creating Model

```{r}
set.seed(10)

sim_regression = function(n = 30, beta0 = 2, beta1) {
  
  sim_data = tibble(
    x = rnorm(n, mean = 0, sd = 1),
    y = beta0 + beta1*x + rnorm(n, mean = 0, sd = sqrt(50)))
  
  ls_fit = lm(y ~ x, data = sim_data)
  
  tibble(
    b1_estimate = broom::tidy(ls_fit)[2, 2] %>% 
      as.numeric(),
    p_value = broom::tidy(ls_fit)[2, 5] %>% 
      as.numeric())
  
}
```

#### Generating 10000 datasets from the model

```{r}
sim_results = tibble(b1 = c(0, 1, 2, 3, 4, 5, 6)) %>% 
  mutate(output_lists = map(.x = b1, ~rerun(10000, sim_regression(beta1 = .x))),
         estimate_dfs = map(output_lists, bind_rows)) %>%
  select(-output_lists) %>%
  unnest(estimate_dfs)
```

#### Plot showing the proportion of times the null was rejected (the power of the test) on the y axis and the true value of β1 on the x axis.

```{r}
sim_results %>% 
  group_by(b1) %>%
  count(reject = p_value < 0.05) %>% 
  mutate(power = n/sum(n)*100) %>% 
  filter(reject == TRUE) %>%
  ggplot(aes(x = b1, y = power)) +
  geom_point() + 
  geom_line() +
  theme_minimal() +
  labs(title = "True β1 and Power") +
  xlab("True value of β1") +
  ylab("Power") +
  theme(plot.title = element_text(hjust = 0.5))
```

In the plot above, as effect size (beta1) increases, the power also increases. 

#### Creating two plots: 1) a plot showing the average estimate of β̂ 1 on the y axis and the true value of β1 on the x axis. 2) plot the average estimate of β̂ 1 only in samples for which the null was rejected on the y axis and the true value of β1 on the x axis. 

```{r}
p1 = 
  sim_results %>% 
  group_by(b1) %>%
  summarise(mean = mean(b1_estimate)) %>% 
  ggplot(aes(x = b1, y = mean)) + 
  geom_point() + 
  geom_line() + 
  labs(title = "Average β1 estimate by true β1") +
  xlab("True β1") +
  ylab("Mean β1 estimate")


p2 = 
  sim_results %>% 
  filter(p_value < 0.05) %>% 
  group_by(b1) %>%
  summarise(mean = mean(b1_estimate)) %>% 
  ggplot(aes(x = b1, y = mean)) + 
  geom_point() +
  geom_line() +
  labs(title = "Average β1 estimate by true β1, subset") +
  xlab("True β1") +
  ylab("Mean β1 estimate")
 
p1 + p2
```

The sample average of the beta1 estimate across tests for which the null is rejected is not approximately  equal to the true value of beta1. The beta1 estimate is higher for true beta1 values that are closer to 0. The difference between the beta1 estimate and the true beta1 value decreases as the true beta1 values increase because as the true beta1 increases, power increases with it. 